---
title: "flow control and functions"
output: html_document
author: "Jurre Hageman"
---

This exercise is about flow control and functions.
Disclaimer: this exercise is for learning the basics of flow control and creating functions.
We recommend that you use built in R functions later on in your code if available.
Built in functions are tested and highly optimized.

# Flow control

Suppose we have some random numbers:


```{r}
set.seed(123) #this line will assure that everebody will get the same result
my_nums <- sample(10000, 1000) #100 random numbers between 1 and 10000
```

## Q1

Use a for loop and flow control to find the largest number in my_nums. 
Do not use the `max` function. Don not use the `sort` function.
Verify your result using the `max` function.

```{r}
#<YOUR CODE HERE>
```


## Q2

Put the code in a function and assign it to the variable `my_max`. 
The function should print `max value: x` with x as the max value.
Test your function for the expected result.

```{r}
#<YOUR CODE HERE>
```

## Q3

Rewrite your function that it returns the max value. Catch the returned value in a variable `max_my_nums`. Print `max_my_nums` to the screen

```{r}
#<YOUR CODE HERE>
```


## Q4

Use flow control to count and report the number of even and odd numbers in `my_nums`.
Do not make use of built in R functions.


```{r}
#<YOUR CODE HERE>
```


## Q5

Put the above code in a function and name it `count_odd_even`.
Print the number of evens and the number of odd values from within the function.
Provide a user friendly message.

```{r}
#<YOUR CODE HERE>
```

## Q6

Get the index in the vector of the highest  number.
Put your code in a function. Name it `index_of_highest`.
The function should return the index of the highest number.

```{r}
#<YOUR CODE HERE>
```

## Q7

Write a function that uses the Pythagorean theorem to calculate the hypotenuse based on the height and base of a right-angled triangle. The function `calc_hypotenuse` should return the hypotenuse. If no height or base is given, use a default value of `1`.
test your function with height = 3 and base = 4. Test your function if no base or height are given.

```{r}
#<YOUR CODE HERE>
```

## Q8

Write a function that will use the quadratic equation (ABC formula) to solve an equation in the form ax^2 + bx + c = 0. Call the function `abc_solver`.
Use flow control to test the following:
- First calculate the discriminant (you should know the equation from your math lessons)
- Test if the discriminant is negative. If so, return a message that tells the user that the equation can not be solved (it has complex roots).
- Then test if the discriminant is positive. If so, there are two solutions for the equation. Return them in a vector.
- Then test if the discriminant is 0. If so, there is one solution for the equation. Return the solution.

Test your function for the following quadratic models:
ax^1 + 2x + 3 (no solution)
ax^3 + 7x + 4 (two intersections)
ax^2 + 4x + 2 (1 intersection)

You can verify your answers at:
https://www.mathsisfun.com/quadratic-equation-solver.html


```{r}
#<YOUR CODE HERE>
```


The last two exercises are about the `ifelse` function.

Suppose we have a data frame with some random values:


```{r}
set.seed(123)
val1 <- 1:100
val2 <- sample(1000, 100)
df <- data.frame(x = val1, y = val2)
head(df)
```


## Q9

Add an column to the df named `even`. If the value in y is even the value in `even` will become `yes` and else `no`.

```{r}
#<YOUR CODE HERE>
```

## Q10

Add a column named `larger_then_mean` to the data frame.
Test for each value in y if it is larger than the mean. If so the value will be `yes` else `no`.

```{r}
#<YOUR CODE HERE>
```


**The end**