---
title: "Apply and aggregate"
output: html_document
author: "Jurre Hageman"
---

This exercise is about the `apply` and `aggregate` functions.


# Apply

There are several `apply` functions.
We will cover:
- apply
- lapply
- sapply

Fist check the documentation of the various apply functions:


```{r}
?apply
```



```{r}
?tapply
```



```{r}
?sapply
```

Let's first create some vectors with some numbers:


```{r}
set.seed(123)

x1 <- rnorm(n = 100, mean = 20, sd = 4)
x2 <- rnorm(n = 100, mean = 40, sd = 4)
x3 <- rnorm(n = 100, mean = 80, sd = 8)
```



## Q1

Create a data frame using the above functions from the 3 vectors. Name the data frame `my_df`.
Use the head function to print the first 10 observations.

```{r}
#<YOUR CODE HERE>
```

## Q2

Calculate the mean for every column. First try to use `mean(my_df)`. What error do you get?
Why would this give an error?

```{r}
#<YOUR CODE HERE>
```

## Q3

Use all different apply functions and compare the results.
Read the documentation for the correct arguments.

regular apply
```{r}
#<YOUR CODE HERE>
```


lapply
```{r}
#<YOUR CODE HERE>
```

sapply
```{r}
#<YOUR CODE HERE>
```

## Q4

Now calculate the mean of each row. What apply function would you use?

```{r}
#<YOUR CODE HERE>
```

## Q5

Make a list of the vectors `x1`, `x2`, `x3`. Name the list `my_list`

```{r}
#<YOUR CODE HERE>
```


## Q6

Use all the apply functions to calculate the mean of the items in the list.
One of the apply functions will not work.
Which one?
Compare the output of the results.

Regular apply
```{r}
#<YOUR CODE HERE>
```

lapply
```{r}
#<YOUR CODE HERE>
```

sapply
```{r}
#<YOUR CODE HERE>
```

## Q7

Mean reduces data.
We will now test the apply function on a non-reducing function.
Use `lapply` to calculate the log with base 2 on all items in the list.
Catch the result in a variable named `my_res_lapply`.
Print the datatype of `my_res_lapply`.
Print the first 10 items for each item in the list using a `for` loop.
Compare the results.

lapply
```{r}
#<YOUR CODE HERE>
```

## Q8

Use `sapply` to calculate the log with base 2 on all items in the list.
Catch the result in a variable named `my_res_sapply`.
Print the datatype of `my_res_sapply`.
Print the first 10 items using the `head` function.
Compare the results.


sapply
```{r}
#<YOUR CODE HERE>
```


# A bigger data set 

We will use the `apply` and `aggregate` functions on a slightly larger data set with fictive gene array data. The text file is located in the folder `files`.
The name of the file is `file1.txt`.

The columns are:
- transcript_ID: NM_ entry number of the reference mRNA.
- fold_induction: the same rules for repression, no change and induction are valid here.
- family: the protein family that the gene encodes for.
- location: the cellular localization of the gene product.

## Q9

Load the file and assign it to the variable `gene_array`.


```{r}
#<YOUR CODE HERE>
```

## Q10

Create an extra column `log2.fold.induction` with a base 2 log transformation of the `fold.induction` column. 

```{r}
#<YOUR CODE HERE>
```

## Q11

Use `apply` to calculate the `mean` for each numeric/integer column in the `gene_array` data frame.
Use the `na.rm` argument as extra argument in apply to deal with missing values correctly.

```{r}
#<YOUR CODE HERE>
```

## Q11

Use `apply` to calculate the `sd` for each numeric column in the `gene_array` data frame.
Use the `na.rm` argument as extra argument in apply to deal with missing values correctly.

## Q12

```{r}
#<YOUR CODE HERE>
```


## Q13

Use the `aggregate` function to investigate the mean log2 induction for each gene family.
Note that aggregate will deal with `NA` values by default. 

```{r}
#<YOUR CODE HERE>
```

## Q14

Use the `aggregate` function to investigate the mean log2 induction for each gene family per cellular compartment.


```{r}
#<YOUR CODE HERE>
```


**The end**