---
title: "Extra material to practice with dataframes and plots"
author: "Jurre Hageman"
date: "May 14, 2020"
---

All data used in this document are real lab-data obtained in the indicated courses.

# Lowry BOVR Theme 6

The Lowry method is a widely used spectroscopic method to quantify protein amounts in solutions.
Here are some real Lowry results for the quantification of proteins:
The calibration curve is made using BSA (bovine serum albumine) as a standard.
The concentrations are in mg/ml:

```{r}
x <- c(0, 0, 0.025, 0.025, 0.075, 0.075, 0.125, 0.125)
x
```

These are the obtained absorbtion values at 750 nm:

```{r}
y <- c(0.063, 0.033, 0.16, 0.181, 0.346, 0.352, 0.491, 0.488)
y
```

## Q1

Add the two vectors to a dataframe with the variable name `dilution`:
Do not add column names yet.

```{r}
# Your code here


```

## Q2

Add the appropriate column names: 
- prot_conc
- abs

```{r}
# Your code here

```

You forgot to include the last datapoints:

```{r}
x2 <- c(0.175,	0.175,	0.25,	0.25)
y2 <- c(0.597,	0.595,	0.743,	0.742)
```

## Q3

Generate a dataframe assigned to the variable `df_temp` from these datapoints. Now directly add the appropriate column names:

```{r}
# Your code here

```

## Q4

Now add the second dataframe `df_temp` to `dilution` and assign it to the dataframe `dilution` (thus overwrite the variable name):

```{r}
# Your code here

```

## Q5

Generate a line plot with the following settings:
- Title: "Absorbance as a function of protein concentration"
- Axis labels with clear indicated quantities and units.
- Datapoints and connector lines should be visible and have a blue color.
- Y-axis lower-limit should be 0, upper-limit should be 1.


```{r}
# Your code here

```

## Q6
The dataframe now contains the duplo values in rows e.g.:

concentration 1 measurement 1, OD value measurement 1
concentration 1 measurement 2, OD value measurement 2
concentration 2 measurement 1, OD value measurement 1
concentration 2 measurement 2, OD value measurement 2

We now will make a dataframe with the duplo values in seperate columns e.g.:

concentration 1 measurement 1, OD value measurement 1, OD value measurement 2
concentration 2 measurement 1, OD value measurement 1, OD value measurement 2

Generate a new dataframe from the data with the duplo measurements side-by-side.
- Fist generate a dataframe `even`using the even numbered rows.
- Then generate a dataframe `odd`using the odd numbered rows.
Hint: use subsetting with TRUE and FALSE and vector recycling.

example:

```{r}
df <- data.frame(1:10, 11:20)
df
df[c(T, F), ]
```


Concatenate these dataframes and assign it to the variable `dilution_duplo`.

```{r}
# Your code here

```

## Q7

The result will be a dataframe with two columns named `prot_conc`.
Delete the second column named `prot_conc`.

```{r}
# Your code here

```

## Q8

Rename the column names to `prot_conc`, `abs1` and `abs2`:

```{r}
# Your code here

```

## Q9

Add a column `mean` and calculate the mean of the two abs measurements:
For this simple example, calculate the mean yourself without using the mean function.
The mean function will not work here but as a challenge you may want to use the function `rowMeans`.


```{r}
# Your code here

```

## Q10

Generate a **bar** plot with the following settings:
- Title: "Absorbance versus protein concentration"
- Axis labels with clear indicated quantities and units.
- Y-axis lower-limit should be 0, upper-limit should be 1. 
- Plot the mean absorbance on the Y-axis.
- Bars should have a green color.

```{r}
# Your code here

```


# HPLC BOVR Theme 7

The data below are from the experiment "Thymol quantification in Thyme".
This data are obtained by a reverse-phase HPLC method using thymol in a standard curve. Peak areas were calculated using an integrator device. Note that peak areas are dimensionless. The concentration thymol is in μg/ml.

- co = concentration
- pa = peak area

```{r}
co <- c(100, 100, 75, 75, 50, 50, 25, 25, 10, 10, 5, 5)
pa <- c(1969017, 1858012, 1399762, 1449423, 963014, 832137, 467856, 562012, 200123, 145545, 94567, 64752)
```

## Q11

Add the vectors to a dataframe named `hplc_data`. Use appropriate and clear column names.

```{r}
# Your code here

```

## Q12

The concentrations are in descending order. Sort the dataframe so that the concentrations are in ascending order.

```{r}
# Your code here

```

## Q13

Generate a scatter plot with the following settings:
- Title: "Peak area as a function of thymol concentration"
- Axis labels with clear indicated quantities and units.
- Datapoints should be visible (no connector lines).
- Datapoints should be blue
- Add a red colored linear regression line with a lineweigth of 2.


```{r}
# Your code here

```


**THE END**