---
title: "Loading data to dataframes"
author: "Jurre Hageman and Michiel Noback"
date: "2020-05-24"
---
# Load data in dataframe

The purpose of this excersise is simple: Load the data in a dataframe using the read.table function.
Check if this is done in an appropriate manner by printing your dataframe.


A test file named `test.txt` is located in the directory `test_dir`. In the next code chunk, we will open the test file using the `read.table` function and assign the contents to the `my_data` variable:


```{r}
my_dir <- "test_dir"
my_file <- "test.txt"
my_path <- paste0(my_dir, "/", my_file)
my_data <- read.table(my_path)
my_data
``` 

The loaded and the field seperator used was a space.
my_data points to a dataframe object consisting of a sinle row and 5 columns in memory:


```{r}
class(my_data)
dim(my_data)
```

# The files:
The files in the `files`directory contains some (fictive) gene-array data. The dataset contains a selection of induced transcripts after some stimulus. The columns represent:
- the mRNA entry
- fold induction after some stimulus
- the protein entry number
- protein length in number of amino acids of the corresponding NP entry
- the protein family the protein belongs to
- the predicted cellular localization

Whatever the contents of a file, you always need to address (some of) these questions:

- Are there comment lines at the top?
- Is there a header line with column names?
- What is the column separator?
- Are there quotes around character data?
- How are missing values encoded?
- How are numeric values encoded?
- What is the type in each column?

Also: read the help for the `read.table` function carefully.

```{r}
?read.table
```

Use the code block below to read the content of the files in the `files` directory in a dataframe object and assign it to the variable name `df`. There are 25 different files to practice your file loading skills.


```{r}
# Replace the NULL with the correct code:
my_dir <- NULL
my_file <- NULL
my_path <- NULL
my_data <- NULL
my_data
```


## file01.txt

```{r}
#<YOUR CODE HERE>
```

## file02.txt

```{r}
#<YOUR CODE HERE>
```

## file03.txt

```{r}
#<YOUR CODE HERE>
```

## file04.txt

```{r}
#<YOUR CODE HERE>
```

## file05.txt

```{r}
#<YOUR CODE HERE>
```

## file06.txt

```{r}
#<YOUR CODE HERE>
```

## file07.txt

```{r}
#<YOUR CODE HERE>
```

## file08.txt

```{r}
#<YOUR CODE HERE>
```

## file09.txt

```{r}
#<YOUR CODE HERE>
```

## file10.txt

```{r}
#<YOUR CODE HERE>
```

## file11.txt

```{r}
#<YOUR CODE HERE>
```

## file12.txt

```{r}
#<YOUR CODE HERE>
```

## file13.txt

```{r}
#<YOUR CODE HERE>
```

## file14.txt

```{r}
#<YOUR CODE HERE>
```

## file15.txt

```{r}
#<YOUR CODE HERE>
```

## file16.txt

```{r}
#<YOUR CODE HERE>
```

## file17.txt

```{r}
#<YOUR CODE HERE>
```

## file18.txt

```{r}
#<YOUR CODE HERE>
```

## file19.txt

```{r}
#<YOUR CODE HERE>
```

## file20.txt

```{r}
#<YOUR CODE HERE>
```

## file21.txt

```{r}
#<YOUR CODE HERE>
```

## file22.txt

```{r}
#<YOUR CODE HERE>
```

## file23.txt

```{r}
#<YOUR CODE HERE>
```

## file24.txt

```{r}
#<YOUR CODE HERE>
```

## file25.txt

```{r}
#<YOUR CODE HERE>
```
